---
title: "Export data from inboveg database: terrestrial habitat types and aquatic habitat types measured by INBO"
date: "`r lubridate::now()`"
output: 
  html_document:
    number_sections: yes
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(odbc)
library(inbodb)
library(git2rdata)
library(n2khab)
```

# Query INBOVEG terrestrial habitat types

## Query database

```{r}
update <- TRUE
```

```{r connection and query, eval= update}
 
connection_inboveg <- connect_inbo_dbase(database_name = "D0010_00_Cydonia")

mhq_terr_surveys <- c("N2000meetnet_Grasland", "N2000meetnet_Moerassen", "N2000meetnet_Grasland_BHM", "N2000meetnet_Moeras_BHM", "N2000meetnet_Duinen_BHM", "N2000_Duinen_Fieldapp", "N2000meetnet_Bos_91E0_sf", "N2000meetnet_Heide_BHM", "N2000meetnet_Heide_BHM_2022", "N2000_6510_droogte", "Schelde-estuarium")

classif_mhq <- get_inboveg_classification(connection = connection_inboveg, survey_name = mhq_terr_surveys, collect = TRUE, multiple = TRUE) %>%
    select(recording_givid = RecordingGivid, survey = Name, type_observed = Classif, classif_type = ActionGroup, classif_key = ListName, type_cover = Cover) %>%
    filter(classif_type == "N2k") %>%
    select(-classif_type)

header_mhq <- get_inboveg_header(connection = connection_inboveg, survey_name = mhq_terr_surveys, collect = TRUE, multiple = TRUE) %>%
    select(recording_givid = RecordingGivid, survey = SurveyName, user_reference = UserReference, area = Area, vague_date_begin = VagueDateBegin, vague_date_end = VagueDateEnd, latitude = Latitude, longitude = Longitude)

recordings_mhq <- get_inboveg_recording(connection = connection_inboveg, survey_name = mhq_terr_surveys, collect = TRUE, multiple = TRUE) %>%
    select(survey = Name, recording_givid = RecordingGivid, layer_code = LayerCode, layer_cover = CoverCode, name_original = OriginalName, name_scientific = ScientificName, phenology_code = PhenologyCode, species_cover_code = CoverageCode, species_cover = PctValue, scale = RecordingScale)

layers_mhq <- get_inboveg_layer_cover(connection = connection_inboveg, survey_name = mhq_terr_surveys, multiple = TRUE) %>%
  select(survey = Name, recording_givid = RecordingGivid, layer_code = LayerCode, layer_description = LayerDescription, layer_cover = CoverCode)

qualifiers_mhq <- get_inboveg_qualifier(connection = connection_inboveg, survey_name = mhq_terr_surveys, qualifier_type = "MQ", multiple = TRUE) %>%
    select(survey = Name, recording_givid = RecordingGivid, user_reference = UserReference, structure_var_code = Q2Code, structure_var = Q2Description, cover = Q3Description, cover_code = Q3Code) %>%
    mutate(structure_var_code = tolower(structure_var_code),
           structure_var = tolower(structure_var),
           cover = as.numeric(cover)) %>%
  mutate(cover = ifelse(is.na(cover), 
                        ifelse(cover_code == "ZS", 0.5, as.numeric(cover_code)),
                        cover))

site_qualifiers_moneos <-  get_inboveg_qualifier(connection = connection_inboveg, survey_name = c("N2000meetnet_Bos_91E0_sf", "Schelde-estuarium"), qualifier_type = "SQ", multiple = TRUE) %>%
  select(survey = Name, recording_givid = RecordingGivid, plot_id = Q1Code)

```

## Coordinates MONEOS

The coordinates from the MONEOS-PQ's are not in INBOVEG (yet?).

For now we store a csv-file with coordinates in [this google drive folder](https://drive.google.com/drive/folders/1tl0gtQrFdmMtgupffkX8uuMJ0uJbeWKL).

```{r}

path <- fileman_up("n2khab-mhq-data")

coordinates_moneos_91E0_sf <- read_csv2(file.path(path, "raw/coordinates_inbo/CoordinatenExtraPQ_LSVI_91E0_2025-01-07.csv")) %>%
  select(plot_id = "PQ-nummer", plot_pl = PQpl, x = LamX, y = LamY, grts_ranking = Ranking) %>%
  filter(!is.na(plot_id)) %>%
  arrange(plot_id)

coordinates_moneos_1330_da <- read_csv2(file.path(path, "raw/coordinates_inbo/coördinaten_1330_da.csv")) %>%
  select(plot_id = id,  x_mean = Lambert_X_mean, y_mean = Lambert_Y_mean, x_l = Lambert_X_l, y_l = Lambert_Y_l, x_r = Lambert_X_r, y_r = Lambert_Y_r) %>%
  arrange(plot_id)
```

## Structure data not in INBOVEG

The structure variables of 1330_da are not in INBOVEG. 

The original data is stored in [this google drive folder](https://drive.google.com/drive/folders/1tl0gtQrFdmMtgupffkX8uuMJ0uJbeWKL).

```{r}
structure_1330_da_wide <- read_csv2(file.path(path, "raw/structure_variables/Structuurgegevens_1330_da.csv"))

structure_1330_da <- structure_1330_da_wide %>%
  gather( starts_with("ZSCPQ"), key = "plot_id", value =  "value") %>%
  mutate(value = ifelse(Voorwaarde %in% c("schorklifvegetaties aanwezig", "habitattype lager dan het klif") & value == "nvt", "nee",
                         value)) %>%
  filter(!is.na(value)) %>%
  filter(Versie != "") %>%
  group_by(plot_id, Voorwaarde) %>%
  summarise(value = unique(value)) %>%
  ungroup() %>%
  filter(!value %in% c("", "?", "nvt")) %>%
  rename(structure_var = Voorwaarde) %>%
  mutate(date = as.Date("2018-11-06"))

```

## Write result to .vc file

```{r, eval= update}

path <- str_c(fileman_up("n2khab-mhq-data"), "/processed")

if (!dir.exists(path)) {
    
  dir.create(path)  
    
}

path <- str_c(fileman_up("n2khab-mhq-data"), "/processed/inboveg_mhq_terr")

if (!dir.exists(path)) {
    
  dir.create(path)  
    
}

write_vc(classif_mhq, "classif_mhq_terr", root = path,  sorting = c("recording_givid", "type_observed"), strict = FALSE)
write_vc(header_mhq, "header_mhq_terr", root = path,  sorting = c("recording_givid"), strict = FALSE)
write_vc(qualifiers_mhq, "structure_mhq_terr", root = path,  sorting = c("recording_givid", "structure_var"), strict = FALSE)
write_vc(layers_mhq, "layers_mhq_terr", root = path,  sorting = c("recording_givid", "layer_code"))
write_vc(recordings_mhq, "vegetation_mhq_terr", root = path,  sorting = c("recording_givid", "layer_code", "name_scientific"))
write_vc(site_qualifiers_moneos, "site_qualifiers_moneos", root = path, sorting = c("recording_givid"))
write_vc(coordinates_moneos_91E0_sf, "coordinates_moneos_91E0_sf", root = path, sorting = c("plot_id", "plot_pl"))
write_vc(coordinates_moneos_1330_da, "coordinates_moneos_1330_da", root = path, sorting = c("plot_id"))
write_vc(structure_1330_da, "structure_mhq_1330_da", root = path, sorting = c("plot_id", "structure_var"), strict = FALSE)
```

# Query INBOVEG aquatic habitat types

## Query database

```{r}
update <- TRUE
```

```{r connection and query aq, eval= update}
 
connection_inboveg <- connect_inbo_dbase(database_name = "D0010_00_Cydonia")
surveys <- c("HT31xx_LSVI_StilstaandeWateren", "HT31xx_Plassen", "HT3260",
             "HT2190_a_LSVI_StilstaandeWateren")

classif_mhq <- get_inboveg_classification(connection = connection_inboveg, 
                                          survey_name = surveys, multiple = TRUE,
                                          collect = TRUE) %>%
    select(recording_givid = RecordingGivid, survey = Name, type_observed = Classif, 
           classif_type = ActionGroup, classif_key = ListName, type_cover = Cover) %>%
    filter(classif_type == "N2k") %>%
    select(-classif_type)

header_mhq <- get_inboveg_header(connection = connection_inboveg, 
                                 survey_name = surveys, multiple = TRUE,
                                 collect = TRUE) %>%
    select(recording_givid = RecordingGivid, survey = SurveyName, user_reference = UserReference, 
           location = LocationCode, area = Area, vague_date_begin = VagueDateBegin, 
           vague_date_end = VagueDateEnd, latitude = Latitude, longitude = Longitude)

recordings_mhq <- get_inboveg_recording(connection = connection_inboveg, 
                                        survey_name = surveys, multiple = TRUE,
                                        collect = TRUE) %>%
    select(survey = Name, recording_givid = RecordingGivid, layer_code = LayerCode, 
           layer_cover = CoverCode, name_original = OriginalName, name_scientific = ScientificName, 
           phenology_code = PhenologyCode, species_cover_code = CoverageCode, species_cover = PctValue, 
           scale = RecordingScale)

qualifiers_m_mhq <- get_inboveg_qualifier(connection = connection_inboveg, 
                                        survey_name = surveys, multiple = TRUE,
                                        qualifier_type = "MQ") %>%
    filter(Q1Code %in% c("A", "B")) %>% # select only present and recent measurements (no winter observations)
    select(survey = Name, recording_givid = RecordingGivid, user_reference = UserReference, 
           var_code = Q2Code, var = Q2Description, 
           var_value_code = Q3Code, var_description_or_value = Q3Description, 
           value = Elucidation) %>%
    mutate(var_code = tolower(var_code), var = tolower(var)) %>% 
    filter(!(value %in% c("NA", "n/a")))  

qualifiers_s_mhq <- get_inboveg_qualifier(connection = connection_inboveg,
                                 survey_name = surveys, multiple = TRUE, 
                                 qualifier_type = "SQ") %>% 
  select(survey = Name, recording_givid = RecordingGivid, user_reference = UserReference, 
         value_code = Q1Code, value = Q1Description, 
         comment = Elucidation)
  
qualifiers_l_mhq <- get_inboveg_layer_cover(connection = connection_inboveg, 
                                       survey_name = surveys,
                                       multiple = TRUE) %>% 
  select(survey = Name, recording_givid = RecordingGivid, user_reference = UserReference,
         var_code = LayerCode, var = LayerDescription, value = CoverCode, value_perc = Percentage)

dbDisconnect(connection_inboveg)
```

## Write result to .vc file

```{r, eval= update}

path <- str_c(fileman_up("n2khab-mhq-data"), "/processed")

if (!dir.exists(path)) {
    
  dir.create(path)  
    
}

path <- str_c(fileman_up("n2khab-mhq-data"), "/processed/inboveg_mhq_aq")

if (!dir.exists(path)) {
    
  dir.create(path)  
    
}

write_vc(classif_mhq, "classif_mhq_aq", root = path,  
         sorting = c("recording_givid", "type_observed"), strict = FALSE)
write_vc(header_mhq, "header_mhq_aq", root = path,  
         sorting = c("recording_givid"), strict = FALSE)
write_vc(qualifiers_m_mhq, "structure_mhq_aq", root = path,  
         sorting = c("recording_givid", "var")) 
write_vc(qualifiers_s_mhq, "sitequal_mhq_aq", root = path,  
         sorting = c("recording_givid")) 
write_vc(qualifiers_l_mhq, "layerqual_mhq_aq", root = path,  
         sorting = c("recording_givid", "var"), strict = FALSE) 
write_vc(recordings_mhq, "vegetation_mhq_aq", root = path,  
         sorting = c("recording_givid", "layer_code", "name_scientific"))
```