---
title: "Export data from the forest inventory analysis database"
date: "`r lubridate::now()`"
output: 
  html_document:
    number_sections: yes
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(git2rdata)
library(n2khab)
library(RODBC)
```

# Analysis database

We make use of following data sources (see [google drive](https://drive.google.com/drive/folders/1TjONf7WuoEYM6TzfHp9UW5KVw6ljMDgs)):

+ `vbi_strata_v2024-01-24.accdb`: observed habitat type (only available for the more recently measured plots)
+ `vbi_meetproces_v2022-11-15.accdb`: recorded coordinates
+ `vbi_analysedatabank_v2024-07-01.accdb`: contains information on the plots for which vegetation and dendrometric data is recorded

```{r}

data_path_vbi <- file.path(fileman_up("n2khab-mhq-data"), "/raw/vbi")

db_strata <- "vbi_strata_v2024-01-24.accdb"
db_meetproces <- "vbi_meetproces_v2022-11-15.accdb"
db_analyse <- "vbi_analysedatabank_v2024-07-01.accdb"

connect_db <-   odbcConnectAccess2007(file.path(data_path_vbi, db_strata))
strata_orig <- sqlFetch(connect_db, "tblplotstratadynamic")
odbcClose(connect_db)

connect_db <-   odbcConnectAccess2007(file.path(data_path_vbi, db_meetproces))
coordinaten_orig <- sqlFetch(connect_db, "tblcoordinaten")
plot_weights_orig <- sqlFetch(connect_db,"tblPlotWeights")
odbcClose(connect_db)

connect_db <-   odbcConnectAccess2007(file.path(data_path_vbi, db_analyse))
bestandskarakt_orig <- sqlFetch(connect_db, "tbl2BestandskaraktKwal")
cover_species_orig <- sqlFetch(connect_db, "tbl11VegetatieSamenstelling")
vegetatie_orig <- sqlFetch(connect_db, "tbl4Vegetatie")
boomsoort_orig <- sqlFetch(connect_db, "tbl10BoomSoorten")
trees_orig <- sqlFetch(connect_db, "tbl0Boom")
deadwood_orig <-  sqlFetch(connect_db, "tbl6DeadWood")
odbcClose(connect_db)

strata <- strata_orig %>%
  select(plot_id = IDPlots, segment_id = IDSegments, periode = Periode, landuse = LandUse, stand_type = StandType, type_observed = Habitat) %>%
  filter(!is.na(stand_type)) %>%
  select(-stand_type)

coordinaten <- coordinaten_orig %>%
  select(plot_id = idplots, x, y, type_coord = type, periode, plotweight)
# 
# coord_recent <- coordinaten %>%
#   semi_join(cover_species_vbi, by = c("plot_id")) %>%
#   group_by(plot_id) %>%
#   slice_max(periode) %>%
#   ungroup()

#write_csv2(coord_recent, "../output/coord_vbi.csv")
```

## Vegetation

```{r}
cover_species_vbi <- cover_species_orig %>%
  select(plot_id = IDPlots, vbi_cycle = Periode, layer = Layer, name_nl = NameNl, 
         name_sc = NameSc, cover_mean = Cover,  cover_description =  Cover_categ, is_tree = Tree) %>%
  mutate(layer = ifelse(layer == "b", "treelayer",
                      ifelse(layer == "k", "herblayer",
                             ifelse(layer == "s", "shrublayer", 
                                    ifelse(layer == "m", "mosslayer", NA)))),
         is_tree = is_tree == 1)

cover_veglayers_vbi <- cover_species_vbi %>%
  group_by(plot_id, vbi_cycle, layer) %>%
  summarise(cover_mean = 1 - prod(1 - cover_mean/100, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(cover_mean = round(cover_mean * 100, 1))
```
## Treesa3a4

```{r}
trees_a3a4 <- trees_orig %>%
  filter(ID != 0) %>%
  mutate(dbh_mm = Diameter_cm * 10,
         coppice_individual = ifelse(Coppice_IndividualCode == 10, "Individuele boom", "Hakhoutstoof"),
         intact_tree = ifelse(IntactTreeCode == 10, "Intacte boom", "Niet intacte boom")) %>%
  select(plot_id = IDPlots, segment_id = IDSegments, tree_id = ID, vbi_cycle = Periode, perimeter_cm = Perimeter_cm, dbh_mm, height_m = Height_m,
         name_nl = NameNl, name_sc = Species, status_tree = StatusTree, coppice_individual, intact_tree, basalarea_m2 = BasalArea_m2, vol_stem_m3 = Volume_m3, vol_tot_m3 = VolumeInclCrown_m3)

```

## Stand description

```{r}
bestandskarakt <- bestandskarakt_orig %>%
  select(plot_id = IDPlots, segment_id = IDSegments, vbi_cycle = Periode, stand_type = v2_StandType, stand_age = v5_StandAge, stand_mixture = v4_MixType) %>%
  pivot_longer(cols = c("stand_type", "stand_age", "stand_mixture"), names_to = "variable", values_to = "category")
```

## Plot weights

```{r}
plot_weight <- trees_orig %>%
  filter(ID != 0) %>%
  select(plot_id = IDPlots, segment_id = IDSegments, vbi_cycle = Periode, plot_weight = PlotWeight, segment_weight = SegmentWeight) %>%
  unique() %>%
  arrange(plot_id, vbi_cycle, segment_id) %>%
  mutate(plot_weight = ifelse(is.na(plot_weight) | plot_weight > 1, 1, plot_weight))

plot_areas <- trees_orig %>%
  filter(ID != 0) %>%
  select(plot_id = IDPlots, segment_id = IDSegments, vbi_cycle = Periode, area_a4_m2 = AreaA4_m2, area_a3_m2 = AreaA3_m2,
         area_a4_m2_plot = AreaA4_m2_Plot, area_a3_m2_plot = AreaA3_m2_Plot) %>%
  filter(!(is.na(area_a4_m2_plot) | is.na(area_a3_m2_plot))) %>%
  unique()

plot_weight <- plot_weight %>%
  left_join(plot_areas, by = c("plot_id", "segment_id", "vbi_cycle")) %>%
  mutate(area_a4_m2_plot = ifelse(is.na(area_a4_m2_plot), max(area_a4_m2_plot, na.rm = TRUE), area_a4_m2_plot),
         area_a3_m2_plot = ifelse(is.na(area_a3_m2_plot), max(area_a3_m2_plot, na.rm = TRUE), area_a3_m2_plot),
         segment_weight = ifelse(is.na(segment_weight), 0, segment_weight))
```

```{r, eval = FALSE}
plot_weight_metadata <- plot_weights_orig %>%
  select(plot_id = idplots, segment_id = idsegments, vbi_cycle = periode, plot_weight = plotweight, segment_weight = segmentweight,
         area_a3_m2 = areaa3_m2, area_a4_m2 = areaa4_m2) %>%
  group_by(plot_id, vbi_cycle) %>%
  mutate(area_a3_m2_plot = sum(area_a3_m2),
         area_a4_m2_plot = sum(area_a4_m2)) %>%
  ungroup()

missing_segment <- plot_weight %>%
  anti_join(plot_weight_metadata, by = c("plot_id", "segment_id", "vbi_cycle"))

check_missing_plot <- plot_weight %>%
  anti_join(plot_weight_metadata, by = c("plot_id", "vbi_cycle"))

plot_weight_complete <- plot_weight_metadata %>%
  bind_rows(missing_segment)
```



## Analyse variables

```{r}
veg_av <- vegetatie_orig %>%
  select(plot_id = IDPlots, segment_id = IDSegments, vbi_cycle = Periode, natural_reg = v0_NV, natural_reg_autocht = v4_NVAutochtPres) %>%
  pivot_longer(cols = c("natural_reg", "natural_reg_autocht"),
               names_to = "variable",
               values_to = "value")

trees_av <- boomsoort_orig %>%
  mutate(status_tree = ifelse(StatusTreeCode == 1, "levend",
                              ifelse(StatusTreeCode == 2, "dood", NA))) %>%
  select(plot_id = IDPlots, segment_id = IDSegments, vbi_cycle = Periode, name_nl = SpeciesGroup, status_tree, stem_density_a2_ha = v6_StamtalA2_ha, basal_area_ha = v2_BasalArea_ha, volume_ha = v3_Volume_ha) %>%
  pivot_longer(cols = c("stem_density_a2_ha", "basal_area_ha", "volume_ha"),
               names_to = "variable",
               values_to = "value")

deadwood <- deadwood_orig %>%
  select(plot_id = IDPlots, segment_id = IDSegments, vbi_cycle = Periode, snags_vol = v1_VolumeSnags, logs_vol = v3_VolumeLogs, alive_vol = TotalVolumeLevend) %>%
  pivot_longer(cols = c("snags_vol", "logs_vol", "alive_vol"),
               names_to = "variable",
               values_to = "value")

vbi_av <- veg_av %>%
  bind_rows(trees_a2, deadwood) %>%
  arrange(plot_id, segment_id, vbi_cycle) %>%
  mutate(value = round(value, 3))

```



# Observed habitat types

The observed habitat types in the vbi plots is based on:

+ habitatmap 2023
+ vbi assessments (only for a part of the vbi plots)
+ manual control (when the assessed habitat type does not correspond with the habitatmap 2023)

A file with an overview of the observed types is created with the code in this [repo](https://github.com/inbo/n2khab-mhq-design/tree/master/mhq_terr_forests/source).

A backup of the file is stored [here](https://drive.google.com/drive/folders/1TjONf7WuoEYM6TzfHp9UW5KVw6ljMDgs)

```{r}
type_observed_orig <- read_csv2(file.path(data_path_vbi, "type_observed_vbi.csv")) 

type_observed <- type_observed_orig %>%
  mutate(type = ifelse(type == "91EO_va", "91E0_va", type)) %>%
  rename(vbi_cycle = periode, type_observed = type)
```
# Parameters for volume calculation

```{r}

db_external_data <- "VBIExterneData.accdb"

  query_tarieven2ing <- "
    SELECT
      tblTariefgroepBoomsoort.ID as tree_species_id
      , tblTariefgroepBoomsoort.Value as name_nl
      , tblTarieven_2ing.Tarief as tarief
      , tblTarieven_2ing.groepNaam as tree_species_group
      , tblTarieven_2ing.a
      , tblTarieven_2ing.b
      , tblTarieven_2ing.c
      , tblTarieven_2ing.d
      , tblTarieven_2ing.e
      , tblTarieven_2ing.f
      , tblTarieven_2ing.g
      , tblTarieven_2ing.Formule_type as formule_type
      FROM tblTariefgroepBoomsoort
      INNER JOIN tblTarieven_2ing ON tblTariefgroepBoomsoort.TariefID = tblTarieven_2ing.groepID
      ;"
  
 query_tarieven1ing <- "
    SELECT tblTariefgroepBoomsoort.ID as tree_species_id
    , tblTariefgroepBoomsoort.Value as name_nl
    , tblTariefgroepBoomsoort.TariefID as tarief_id
    , tblTarieven_1ing.groepNaam as tree_species_group
    , tblTarieven_1ing.a
    , tblTarieven_1ing.b
    , tblTarieven_1ing.c
    , tblTarieven_1ing.d
    , tblTarieven_1ing.Tarief as tarief
    FROM tblTariefgroepBoomsoort
    LEFT JOIN tblTarieven_1ing ON tblTariefgroepBoomsoort.TariefID = tblTarieven_1ing.groepID
    ;"

  connect_db <- odbcConnectAccess2007(file.path(data_path_vbi, db_external_data))

  vol_parameters_2input <- sqlQuery(connect_db, query_tarieven2ing, stringsAsFactors = TRUE)
  vol_parameters_1input <- sqlQuery(connect_db, query_tarieven1ing, stringsAsFactors = TRUE)
  
  odbcClose(connect_db)

```



# Write result to .vc file

```{r, eval= update}

path <- str_c(fileman_up("n2khab-mhq-data"), "/processed")

if (!dir.exists(path)) {
    
  dir.create(path)  
    
}

path <- str_c(fileman_up("n2khab-mhq-data"), "/processed/dwh_vbi")

if (!dir.exists(path)) {
    
  dir.create(path)  
    
}

write_vc(cover_species_vbi, "cover_species", root = path, sorting = c("plot_id", "vbi_cycle", "layer", "name_sc"))
write_vc(cover_veglayers_vbi, "cover_veglayers", root = path, sorting = c("plot_id", "vbi_cycle", "layer"))
write_vc(trees_a3a4, "trees_a3a4", root = path, sorting = c("plot_id", "vbi_cycle", "segment_id", "tree_id"), strict = FALSE)
write_vc(type_observed, "type_observed", root = path, sorting = c("plot_id", "vbi_cycle", "type_observed"), strict = FALSE)
write_vc(vbi_av, "analysis_variable", root = path, sorting = c("plot_id", "segment_id", "vbi_cycle", "variable"), strict = FALSE)
write_vc(bestandskarakt, "stand_description", root = path, sorting = c("plot_id", "segment_id", "vbi_cycle", "variable"), strict = FALSE)
write_vc(plot_weight, "plot_weight", root = path, sorting = c("plot_id", "segment_id", "vbi_cycle"), strict = FALSE)
write_vc(vol_parameters_2input, "vol_parameters_2input", root = path, sorting = c("tree_species_id"), strict = FALSE)
write_vc(vol_parameters_1input, "vol_parameters_1input", root = path, sorting = c("tree_species_id"), strict = FALSE)
```
